---
title: "rra - Visualisierung history files"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Set working directory and load files

First set the working directory. Then load all the .csv files into data frames.

```{r}
if (.Platform$OS.type == "windows") {
  strWDpath <- "E:/Paper_1/2_work"
} else {
  strWDpath <- "/home/khochholzer/NAS/Paper_1/2_work"
}
strWDpath <- normalizePath(path = strWDpath, winslash = "/")

# working directory only applies during this code chunk!
source("support_functions.R")

### Complexity ###
# ConvNext
l_cphist_cnB <- read.csv(file.path(get_newest_directory("../output/complexity/convnext_base"), "history.csv"), sep = ",", dec = ".")
l_cphist_cnL <- read.csv(file.path(get_newest_directory("../output/complexity/convnext_large"), "history.csv"), sep = ",", dec = ".")
l_cphist_cnS <- read.csv(file.path(get_newest_directory("../output/complexity/convnext_small"), "history.csv"), sep = ",", dec = ".")
l_cphist_cnT <- read.csv(file.path(get_newest_directory("../output/complexity/convnext_tiny"), "history.csv"), sep = ",", dec = ".")
l_cphist_cnXL <- read.csv(file.path(get_newest_directory("../output/complexity/convnext_xlarge"), "history.csv"), sep = ",", dec = ".")

# EfficientNet
l_cphist_B0 <- read.csv(file.path(get_newest_directory("../output/complexity/efficientnetb0"), "history.csv"), sep = ",", dec = ".")
l_cphist_B2 <- read.csv(file.path(get_newest_directory("../output/complexity/efficientnetb2"), "history.csv"), sep = ",", dec = ".")
l_cphist_B4 <- read.csv(file.path(get_newest_directory("../output/complexity/efficientnetb4"), "history.csv"), sep = ",", dec = ".")
l_cphist_B5 <- read.csv(file.path(get_newest_directory("../output/complexity/efficientnetb5"), "history.csv"), sep = ",", dec = ".")
# Xception
l_cphist_XC <- read.csv(file.path(get_newest_directory("../output/complexity/xception"), "history.csv"), sep = ",", dec = ".")

# Groundcover
l_gchist_cnB <- read.csv(file.path(get_newest_directory("../output/groundcover/convnext_base"), "history.csv"), sep = ",", dec = ".")
l_gchist_cnL <- read.csv(file.path(get_newest_directory("../output/groundcover/convnext_large"), "history.csv"), sep = ",", dec = ".")
l_gchist_cnS <- read.csv(file.path(get_newest_directory("../output/groundcover/convnext_small"), "history.csv"), sep = ",", dec = ".")
l_gchist_cnT <- read.csv(file.path(get_newest_directory("../output/groundcover/convnext_tiny"), "history.csv"), sep = ",", dec = ".")
l_gchist_cnXL <- read.csv(file.path(get_newest_directory("../output/groundcover/convnext_xlarge"), "history.csv"), sep = ",", dec = ".")

l_gchist_B0 <- read.csv(file.path(get_newest_directory("../output/groundcover/efficientnetb0"), "history.csv"), sep = ",", dec = ".")
l_gchist_B2 <- read.csv(file.path(get_newest_directory("../output/groundcover/efficientnetb2"), "history.csv"), sep = ",", dec = ".")
l_gchist_B4 <- read.csv(file.path(get_newest_directory("../output/groundcover/efficientnetb4"), "history.csv"), sep = ",", dec = ".")
l_gchist_B5 <- read.csv(file.path(get_newest_directory("../output/groundcover/efficientnetb5"), "history.csv"), sep = ",", dec = ".")
l_gchist_XC <- read.csv(file.path(get_newest_directory("../output/groundcover/xception"), "history.csv"), sep = ",", dec = ".")
```

# Plot graphs stored in history files

Then go on to plot all the graphs in the history file per metric. For this load ggplot and ggarrange.

```{r}
library(ggpubr)
library(tidyverse)
library(flextable)
```

Function to plot all training and validation metrics per model

```{r}
f_plotMetrics <- function(p_dfCompl, p_dfGC) {
  l_col <- c("Training" = "lightblue", "Validation" = "orange")
  
  r_plot <- ggarrange(
    ggarrange(
      ggplot(p_dfCompl, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1), y = loss)) + geom_point() + geom_line() +
       labs(x = "epoch", title = "complexity") + theme_bw(),
      
      ggplot(p_dfCompl, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1))) +
       geom_line(aes(y = kullback_leibler_divergence, color = "Training"))  +
       geom_line(aes(y = val_kullback_leibler_divergence, color = "Validation")) +
       labs(x = "epoch", color = "Legend:") +
       scale_color_manual(values = l_col) + theme_bw(),
     
      ggplot(p_dfCompl, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1))) +
        geom_line(aes(y = mean_relative_error, color = "Training")) +
        geom_line(aes(y = val_mean_relative_error, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      ggplot(p_dfCompl, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1))) +
        geom_line(aes(y = mean_squared_error, color = "Training")) +
        geom_line(aes(y = val_mean_squared_error, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      ggplot(p_dfCompl, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1))) +
        geom_line(aes(y = accuracy, color = "Training")) +
        geom_line(aes(y = val_accuracy, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      nrow = 3, ncol = 2, common.legend = TRUE, legend = "bottom"
    ),
    
    ggarrange(
      ggplot(p_dfGC, aes(x = seq(from = 1, to = nrow(p_dfGC), by = 1), y = loss)) + geom_point() + geom_line() +
        labs(x = "epoch", title = "groundcover") + theme_bw(),
      
      ggplot(p_dfGC, aes(x = seq(from = 1, to = nrow(p_dfGC), by = 1))) +
        geom_line(aes(y = kullback_leibler_divergence, color = "Training"))  +
        geom_line(aes(y = val_kullback_leibler_divergence, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      ggplot(p_dfGC, aes(x = seq(from = 1, to = nrow(p_dfGC), by = 1))) +
        geom_line(aes(y = mean_relative_error, color = "Training")) +
        geom_line(aes(y = val_mean_relative_error, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      ggplot(p_dfGC, aes(x = seq(from = 1, to = nrow(p_dfGC), by = 1))) +
        geom_line(aes(y = mean_squared_error, color = "Training")) +
        geom_line(aes(y = val_mean_squared_error, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      ggplot(p_dfGC, aes(x = seq(from = 1, to = nrow(p_dfGC), by = 1))) +
        geom_line(aes(y = accuracy, color = "Training")) +
        geom_line(aes(y = val_accuracy, color = "Validation")) +
        labs(x = "epoch", color = "Legend:") +
        scale_color_manual(values = l_col) + theme_bw(),
      
      nrow = 3, ncol = 2, common.legend = TRUE, legend = "bottom"
    )
  )
  return(r_plot)
}
```

## EfficientNet - models

First for EfficientNet - models

### B0

```{r}
l_dfCP <- l_cphist_B0
l_dfGC <- l_gchist_B0

f_plotMetrics(l_cphist_B0, l_gchist_B0)

l_dfCP %>% slice(n())
l_dfGC %>% slice(n())
```

B0 the training and validation runs do not look good, so this model will be excluded in the next steps.

### B2

```{r}
l_dfCP <- l_cphist_B2
l_dfGC <- l_gchist_B2

f_plotMetrics(l_cphist_B2, l_gchist_B2)
```

B2 training and validation runs look better than the B0 ones. The complexity graphs are a bit more "unstable" than the groundcover ones.

### B4

```{r}
l_dfCP <- l_cphist_B4
l_dfGC <- l_gchist_B4

f_plotMetrics(l_cphist_B4, l_gchist_B4)
```

The training and validation graphs do not look good, B4 will also be excluded from the next steps.

### B5

```{r}
l_dfCP <- l_cphist_B5
l_dfGC <- l_gchist_B5

f_plotMetrics(l_cphist_B5, l_gchist_B5)
```

The B5 graphs for complexity look realy good! Groundcover graphs are still okay-isch.

## Xception

Now for the **Xception** model:

```{r}
l_dfCP <- l_cphist_XC
l_dfGC <- l_gchist_XC

f_plotMetrics(l_cphist_XC, l_gchist_XC)

l_col <- c("Training" = "lightblue", "Validation" = "orange")
ggplot(l_dfCP, aes(x = seq(from = 1, to = nrow(l_dfCP), by = 1))) +
  geom_line(aes(y = loss, color = "Training")) +
  geom_line(aes(y = val_loss, color = "Validation")) +
  labs(x = "epoch", color = "Legend") +
  scale_color_manual(values = l_col) + theme_bw()

ggplot(l_dfGC, aes(x = seq(from = 1, to = nrow(l_dfGC), by = 1))) +
  geom_line(aes(y = loss, color = "Training")) +
  geom_line(aes(y = val_loss, color = "Validation")) +
  labs(x = "epoch", color = "Legend") +
  scale_color_manual(values = l_col) + theme_bw()
```

The xception model results look good for complexity and groundcover.

## ConvNext - models

Lastly the new ConvNext models:

### Tiny

```{r}
l_dfCP <- l_cphist_cnT
l_dfGC <- l_gchist_cnT

f_plotMetrics(l_cphist_cnT, l_gchist_cnT)
```

while the loss function is looking good, the validation graphs do not look as good.

### Small

```{r}
l_dfCP <- l_cphist_cnS
l_dfGC <- l_gchist_cnS

f_plotMetrics(l_cphist_cnS, l_gchist_cnS)
```

The graphs from ConvNextSmall look good.

### Base

```{r}
l_dfCP <- l_cphist_cnB
l_dfGC <- l_gchist_cnB

f_plotMetrics(l_cphist_cnB, l_gchist_cnB)
```

The graphs from ConvNextBase also look good.

### Large

```{r}
l_dfCP <- l_cphist_cnL
l_dfGC <- l_gchist_cnL

f_plotMetrics(l_cphist_cnL, l_gchist_cnL)
```

While the graphs for complexity look very good with ConvNextLarge, the groundcover ones do not look as good.

### XLarge

```{r}
l_dfCP <- l_cphist_cnXL
l_dfGC <- l_gchist_cnXL

f_plotMetrics(l_cphist_cnXL, l_gchist_cnXL)
```

# Comparison

## Training Graphs

Here the graphs for ConvNextXLarge also look ok-isch.

| Model              | Complexity | Groundcover | Used in next step |
|--------------------|------------|-------------|-------------------|
| EfficientNet**B0** | \-         | \-          | no                |
| EfficientNet**B2** | \+         | \+          | yes               |
| EfficientNet**B4** | \-         | \-          | no                |
| EfficientNet**B5** | \+         | \~+         | yes               |
| **Xception**       | \+         | \+          | yes               |
| ConvNext**Tiny**   | \~+        | \~+         | yes               |
| ConvNext**Small**  | \+         | \~-         | no                |
| ConvNext**Base**   | \+         | \-          | no                |
| ConvNext**Large**  | \+         | \~-         | no                |
| ConvNext**XLarge** | \~+        | \~-         | no                |

## Training and validation metrics

Just to have a more numeric approach, a table with the last values of training and validation metrics is created. IMPORTEND is result of 10 fold cross-validation (gives robust numbers for the training/validation process) + out of sample test result.

```{r}
l_cphist_B0$model <- "B0"
l_cphist_B2$model <- "B2"
l_cphist_B4$model <- "B4"
l_cphist_B5$model <- "B5"
l_cphist_cnT$model <- "T"
l_cphist_cnS$model <- "S"
l_cphist_cnB$model <- "B"
l_cphist_cnL$model <- "L"
l_cphist_cnXL$model <- "XL"
l_cphist_XC$model <- "XC"

l_gchist_B0$model <- "B0"
l_gchist_B2$model <- "B2"
l_gchist_B4$model <- "B4"
l_gchist_B5$model <- "B5"
l_gchist_cnT$model <- "T"
l_gchist_cnS$model <- "S"
l_gchist_cnB$model <- "B"
l_gchist_cnL$model <- "L"
l_gchist_cnXL$model <- "XL"
l_gchist_XC$model <- "XC"
```

```{r}
# first complexity of surroundings
l_dfCP <- rbind(l_cphist_B0 %>% slice(n()),
                l_cphist_B2 %>% slice(n()),
                l_cphist_B4 %>% slice(n()),
                l_cphist_B5 %>% slice(n()),
                l_cphist_cnT %>% slice(n()),
                l_cphist_cnS %>% slice(n()),
                l_cphist_cnB %>% slice(n()),
                l_cphist_cnL %>% slice(n()),
                l_cphist_cnXL %>% slice(n()),
                l_cphist_XC %>% slice(n())
                )
l_dfCP <- l_dfCP %>% relocate(model)
l_dfCP

# second groundcover
l_dfGC <- rbind(l_gchist_B0 %>% slice(n()),
                l_gchist_B2 %>% slice(n()),
                l_gchist_B4 %>% slice(n()),
                l_gchist_B5 %>% slice(n()),
                l_gchist_cnT %>% slice(n()),
                l_gchist_cnS %>% slice(n()),
                l_gchist_cnB %>% slice(n()),
                l_gchist_cnL %>% slice(n()),
                l_gchist_cnXL %>% slice(n()),
                l_gchist_XC %>% slice(n())
                )
l_dfGC <- l_dfGC %>% relocate(model)
l_dfGC

```
## Train, val metrics
Table 1: Metrics from all DNN model training of context and groundcover photos. Models beginning with B are from the EfficientNet family (e.g. EfficientNetB0). T, S, B, L, XL are abbreviations for the ConvNext families (ConvNextTiny, Small, Base, Large, XLarge respectively). XC is the abbreviation for the Xception model. KL-Divergence represents Kullback-Leibler Divergence, MRE mean relative error, MSE mean squared error.
```{r train_val_metrics}
df_tbl <- bind_rows(
  l_dfCP %>% 
    select(!lr) %>% 
    pivot_longer(
      cols = -model,
      # .value tells R to keep that part of the name as a column header
      names_to = c("type", ".value"),
      # regex captures the prefix (if exists) and the metrics name
      names_pattern = "(?:(val)_)?(.+)"
    ) %>% 
    mutate(
      type = if_else(type == "val", "validation", "training"),
      img_type = "context"
    ),
  l_dfGC %>% 
    select(!lr) %>% 
    pivot_longer(
      cols = -model,
      names_to = c("type", ".value"),
      names_pattern = "(?:(val)_)?(.+)"
    ) %>% 
    mutate(
      type = if_else(type == "val", "validation", "training"),
      img_type = "groundcover"
    )
) %>% 
  select(
    `Image type` = img_type,
    Stage = type,
    Model = model,
    Loss = loss,
    `KL-Divergence` = kullback_leibler_divergence,
    MRE = mean_relative_error,
    MSE = mean_squared_error,
    Accuracy = accuracy
  ) %>% 
  arrange(`Image type`, Stage) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)))

ft <- df_tbl %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(df_tbl), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(df_tbl), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::hline(i = c(4,9,10,14,19,20,24,29,30,34,39)) %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")
ft
ft <- fit_to_width(ft, max_width = 6.5) # for Word file
# save_as_docx(ft, path = "output/sup_tbl1.docx")
# write_csv(df_tbl, file = "output/sup_tbl1.csv")
```

## 10 fold cross-val report
Groundcover --
```{r}
cv_cnt <- read_csv(file = file.path(get_newest_directory("../output/groundcover/convnext_tiny"), "cross_val_res.csv"))
cv_cns <- read_csv(file = file.path(get_newest_directory("../output/groundcover/convnext_small"), "cross_val_res.csv"))
cv_cnb <- read_csv(file = file.path(get_newest_directory("../output/groundcover/convnext_base"), "cross_val_res.csv"))
cv_cnl <- read_csv(file = file.path(get_newest_directory("../output/groundcover/convnext_large"), "cross_val_res.csv"))
cv_cnxl <- read_csv(file = file.path(get_newest_directory("../output/groundcover/convnext_xlarge"), "cross_val_res.csv"))
cv_xc <- read_csv(file = file.path(get_newest_directory("../output/groundcover/xception"), "cross_val_res.csv"))
cv_b0 <- read_csv(file = file.path(get_newest_directory("../output/groundcover/efficientnetb0"), "cross_val_res.csv"))
cv_b2 <- read_csv(file = file.path(get_newest_directory("../output/groundcover/efficientnetb2"), "cross_val_res.csv"))
cv_b4 <- read_csv(file = file.path(get_newest_directory("../output/groundcover/efficientnetb4"), "cross_val_res.csv"))
cv_b5 <- read_csv(file = file.path(get_newest_directory("../output/groundcover/efficientnetb5"), "cross_val_res.csv"))

cv_cnt$model <- "ConvNeXtTiny"
cv_cns$model <- "ConvNeXtSmall"
cv_cnb$model <- "ConvNeXtBase"
cv_cnl$model <- "ConvNeXtLarge"
cv_cnxl$model <- "ConvNeXtXLarge"
cv_xc$model <- "Xception"
cv_b0$model <- "EfficientNetB0"
cv_b2$model <- "EfficientNetB2"
cv_b4$model <- "EfficientNetB4"
cv_b5$model <- "EfficientNetB5"

gc_cv <- bind_rows(
  cv_cnt,
  cv_cns,
  cv_cnb,
  cv_cnl,
  cv_cnxl,
  cv_xc,
  cv_b0,
  cv_b2,
  cv_b4,
  cv_b5
) %>% group_by(model) %>% 
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(-Fold)
gc_cv
```
Complexity --
```{r}
cv_cnt <- read_csv(file = file.path(get_newest_directory("../output/complexity/convnext_tiny"), "cross_val_res.csv"))
cv_cns <- read_csv(file = file.path(get_newest_directory("../output/complexity/convnext_small"), "cross_val_res.csv"))
cv_cnb <- read_csv(file = file.path(get_newest_directory("../output/complexity/convnext_base"), "cross_val_res.csv"))
cv_cnl <- read_csv(file = file.path(get_newest_directory("../output/complexity/convnext_large"), "cross_val_res.csv"))
cv_cnxl <- read_csv(file = file.path(get_newest_directory("../output/complexity/convnext_xlarge"), "cross_val_res.csv"))
cv_xc <- read_csv(file = file.path(get_newest_directory("../output/complexity/xception"), "cross_val_res.csv"))
cv_b0 <- read_csv(file = file.path(get_newest_directory("../output/complexity/efficientnetb0"), "cross_val_res.csv"))
cv_b2 <- read_csv(file = file.path(get_newest_directory("../output/complexity/efficientnetb2"), "cross_val_res.csv"))
cv_b4 <- read_csv(file = file.path(get_newest_directory("../output/complexity/efficientnetb4"), "cross_val_res.csv"))
cv_b5 <- read_csv(file = file.path(get_newest_directory("../output/complexity/efficientnetb5"), "cross_val_res.csv"))

cv_cnt$model <- "ConvNeXtTiny"
cv_cns$model <- "ConvNeXtSmall"
cv_cnb$model <- "ConvNeXtBase"
cv_cnl$model <- "ConvNeXtLarge"
cv_cnxl$model <- "ConvNeXtXLarge"
cv_xc$model <- "Xception"
cv_b0$model <- "EfficientNetB0"
cv_b2$model <- "EfficientNetB2"
cv_b4$model <- "EfficientNetB4"
cv_b5$model <- "EfficientNetB5"

co_cv <- bind_rows(
  cv_cnt,
  cv_cns,
  cv_cnb,
  cv_cnl,
  cv_cnxl,
  cv_xc,
  cv_b0,
  cv_b2,
  cv_b4,
  cv_b5
) %>% group_by(model) %>% 
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(-Fold)
co_cv
```
### Table S3 creation
```{r}
gc_cv$task <- "Groundcover estimation"
co_cv$task <- "Complexity estimation"
cv_df <- bind_rows(gc_cv, co_cv) %>%
  select(-loss) %>%
  relocate(task, .after = model) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3))) %>% 
  mutate(RMSE = round(sqrt(mean_squared_error), digits = 3))

cv_df <- cv_df %>% select(
  Model = model,
  Task = task,
  `KL Divergence` = kullback_leibler_divergence,
  MRE = mean_relative_error,
  MSE = mean_squared_error,
  Accuracy = accuracy,
  RMSE
)

ft <- cv_df %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(cv_df), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(cv_df), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::hline(i = c(5,9,10,15,19)) %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")
ft
ft <- fit_to_width(ft, max_width = 6.5) # for Word file
save_as_docx(ft, path = "output/sup_tbl_crossval.docx")
```

## Test metrics (Table S2)
Supplement table S2. Metrics from out of sample test.
```{r}
# Load test_res.txt for each model
DNN_models <- c(
  "convnext_tiny", "convnext_small", "convnext_base", "convnext_large",
  "convnext_xlarge", "efficientnetb0", "efficientnetb2", "efficientnetb4",
  "efficientnetb5", "xception"
)
# Complexity
dfco_tst <- tibble()
dfgc_tst <- tibble()
for (model in DNN_models) {
  # Complexity
  curr_tblco <- read_csv(
    file = file.path(
      normalizePath(
        path = get_newest_directory(file.path(strWDpath, "output/complexity/", model)),
        winslash = "/"
      ),
      "test_res.txt"
    )
  ) %>% mutate(model = model)
  
  # Groundcover
  curr_tblgc <- read_csv(
    file = file.path(
      normalizePath(
        path = get_newest_directory(file.path(strWDpath, "output/groundcover/", model)),
        winslash = "/"
      ),
      "test_res.txt"
    )
  ) %>% mutate(model = model)
  
  dfco_tst <- bind_rows(dfco_tst, curr_tblco)
  dfgc_tst <- bind_rows(dfgc_tst, curr_tblgc)
}
# Complexity
dfco_tst <- dfco_tst %>% 
  select(
    Model = model,
    Loss = `['loss'`,
    `KL-divergence` = `'kullback_leibler_divergence'`,
    MRE = `'mean_relative_error'`,
    MSE = `'mean_squared_error'`,
    Accuracy = `'accuracy']`
  ) %>% 
  mutate(across(c(Loss, Accuracy), ~str_remove_all(., "\\[|\\]"))) %>% 
  mutate(across(where(is.character), type.convert, as.is = TRUE)) %>% 
  mutate(RMSE = sqrt(MSE)) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)))

# Groundcover
dfgc_tst <- dfgc_tst %>% 
  select(
    Model = model,
    Loss = `['loss'`,
    `KL-divergence` = `'kullback_leibler_divergence'`,
    MRE = `'mean_relative_error'`,
    MSE = `'mean_squared_error'`,
    Accuracy = `'accuracy']`
  ) %>% 
  mutate(across(c(Loss, Accuracy), ~str_remove_all(., "\\[|\\]"))) %>% 
  mutate(across(where(is.character), type.convert, as.is = TRUE)) %>% 
  mutate(RMSE = sqrt(MSE)) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)))

dftst <- bind_rows(
  dfco_tst %>% mutate(Type = "complexity"),
  dfgc_tst %>% mutate(Type = "groundcover")
) %>% 
  select(-Loss)

dftst <- dftst %>% select(
  Model, Type, `KL Divergence` = `KL-divergence`, MRE, MSE, Accuracy, RMSE
)
dftst <- dftst %>% 
  mutate(
    Model = recode(
      Model,
      "efficientnetb0" = "EfficientNetB0",
      "efficientnetb2" = "EfficientNetB2",
      "efficientnetb4" = "EfficientNetB4",
      "efficientnetb5" = "EfficientNetB5",
      "convnext_tiny" = "ConvNeXtTiny",
      "convnext_small" = "ConvNeXtSmall",
      "convnext_base" = "ConvNeXtBase",
      "convnext_large" = "ConvNeXtLarge",
      "convnext_xlarge" = "ConvNeXtXLarge",
      "xception" = "Xception"
    ),
    Type = recode(
      Type,
      "complexity" = "Complexity estimation",
      "groundcover" = "Groundcover estimation"
    )
  )


ft <- dftst %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(dftst), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(dftst), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::align(align = "center", part = "all") %>%
    flextable::hline(i = c(5,9,10,15,19)) %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")

ft <- fit_to_width(ft, max_width = 6.5)
ft
save_as_docx(ft, path = "output/sup_tbl_dnntst.docx")
```


4 Models will be used in the next step: B5, Xception, ConvNextSmall and -Base.

# Cleanup

Remove all variables and dataframes + functions from R environment.

```{r}
rm(
  l_cphist_cnB,
  l_cphist_cnL,
  l_cphist_cnS,
  l_cphist_cnT,
  l_cphist_cnXL,
  l_cphist_B0,
  l_cphist_B2,
  l_cphist_B4,
  l_cphist_B5,
  l_cphist_XC,
  l_gchist_cnB,
  l_gchist_cnL,
  l_gchist_cnS,
  l_gchist_cnT,
  l_gchist_cnXL,
  l_gchist_B0,
  l_gchist_B2,
  l_gchist_B4,
  l_gchist_B5,
  l_gchist_XC
)
rm(ft)
rm(l_col)
rm(l_dfCP, l_dfGC)
rm(f_plotMetrics)
rm(get_newest_directory)
rm(df_tbl)

rm(curr_tblco, curr_tblgc)
rm(dfco_tst, dfgc_tst, dftst)

rm(DNN_models, model, strWDpath)
rm(co_cv, gc_cv, cv_b0, cv_b2, cv_b4, cv_b5)
rm(cv_cnt, cv_cns, cv_cnb, cv_cnl, cv_cnxl)
rm(cv_df, cv_xc)
```
