---
title: "Analysis_Paper1"
author: "Kilian Hochholzer"
date: "2025-02-14"
output: html_document
---

# Content

1.  Technical validation of DNNs
2.  Disturbance detection
3.  Disturbance severity
4.  Post-disturbance development

## Libraries

Loading all needed libraries into the environment & seed for reproductability.

```{r library}
set.seed(42)

library(tidyverse)
library(factoextra) # PCA

library(ggpubr)
library(ggcorrplot)
library(ggforce)
library(progress)

# Machine learning
library(caret)
library(randomForest)

# Colours for plot
library(paletteer)
library(khroma)

library(flextable)
```

## Loading .csv-files

4 different datasets are loaded:

-   Field Data - data gathered in the field with the goal to estimate the development trajectory. Here the trajectory of the 80 disturbed sites in Franconia is stored.

-   BA severity - disturbance severity based on field data. Severity is defined as loss of basal area, between reference site and disturbed site.

-   Groundtruth - "true" value of the pictures (complexity and groundcover)

-   Predictions - data created by Neural Network (DNN)

```{r load_data}
#### Path handling for different OS ####
if (.Platform$OS.type == "windows") {
  strWDpath <- "E:/Paper_1/2_work/"
} else {
  strWDpath <- "/home/khochholzer/NAS/Paper_1/2_work/"
}
strWDpath <- normalizePath(strWDpath)

source("support_functions.R")

# Set path to DNN model directory
strCoDNNPath <- normalizePath(
  path = get_newest_directory(file.path(strWDpath, "output/complexity/xception")),
  winslash = "/"
)
strGcDNNPath <- normalizePath(
  path = get_newest_directory(file.path(strWDpath, "output/groundcover/xception")),
  winslash = "/"
)

#### Load Groundtruth ####
dfCO_og <- read.csv(
  "../../1_dataRaw/compl_cleaned.csv",
  sep = ",",
  dec = "."
)
dfGC_og <- read.csv(
  "../../1_dataRaw/paths_and_labels.csv",
  sep = ",",
  dec = "."
)

#### Load FieldData ####
dfFieldData <- read.csv(
  file = "../reorg_full.csv",
  sep = ",",
  dec = "."
)

# set datatypes
dfFieldData$trip_n  <- as.factor(dfFieldData$trip_n)
dfFieldData$manag   <- as.factor(dfFieldData$manag)
dfFieldData$species <- as.factor(dfFieldData$species)
levels(dfFieldData$manag) <- c("cleared","dead","living")


#### Load newly labeled complexity data ####
dfCO_KH <- readxl::read_excel(
  path = "../../1_dataRaw/compl_KH.xlsx",
  sheet = 1
) %>% filter(!is.na(Baumschicht)) %>% 
  select(
    Img,
    grade_treespecies = Baumschicht,
    grade_shrubs   = Strauchschicht,
    grade_herbs    = Krautschicht,
    grade_grass    = Grassschicht,
    grade_moss     = Moosschicht,
    grade_deadwood = Totholz,
    grade_layers   = Schichtigkeit,
    grade_mixing   = Durchmischung,
    grade_stand_density  = Bestandesdichte
  ) %>% 
  separate_wider_delim(
    Img,
    delim = "_",
    names = c("triplet", "species", "manag", "sub_plt", "orient")
  )
dfCO_KH[dfCO_KH$grade_deadwood > 5,]$grade_deadwood <- 5


#### Process Groundtruth ####
# some data processing (splitting image path into triplet information)
dfCO_og$image_path <- tools::file_path_sans_ext(basename(dfCO_og$image_path))
dfGC_og$image_path <- tools::file_path_sans_ext(basename(dfGC_og$image_path))
dfCO_og <- dfCO_og %>%
  separate_wider_delim(
    image_path, delim = "_",
    names = c("triplet","species","manag","sub_plt","orient")
  )
dfGC_og <- dfGC_og %>%
  separate_wider_delim(
    image_path, delim = "_",
    names = c("triplet","species","manag","sub_plt","orient")
  )

# set columns to datatype factor
dfCO_og$triplet <- as.factor(dfCO_og$triplet)
dfCO_og$species <- as.factor(dfCO_og$species)
dfCO_og$manag   <- as.factor(dfCO_og$manag)
dfGC_og$triplet <- as.factor(dfGC_og$triplet)
dfGC_og$species <- as.factor(dfGC_og$species)
dfGC_og$manag   <- as.factor(dfGC_og$manag)

#### Aggregation of Groundtruth by calculating mean values ####
dfCO_og_agg <- dfCO_og %>%
  select(-sub_plt, -orient) %>% 
  group_by(triplet, manag, species) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  ungroup()
dfGC_og_agg <- dfGC_og %>%
  select(-sub_plt, -orient) %>% 
  group_by(triplet, manag, species) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  ungroup()

# add living to the factor levels
levels(dfCO_og_agg$manag) <- c("cleared","dead","living")
levels(dfGC_og_agg$manag) <- c("cleared","dead","living")


###### Load & process PREDICTION data ######
setwd(file.path(strWDpath, "output"))

dfGC <- read.csv(
  file = file.path(strGcDNNPath, "predictions.csv"),
  sep = ",",
  dec = "."
)
dfCO <- read.csv(
  file = file.path(strCoDNNPath, "predictions.csv"),
  sep = ",",
  dec = "."
)
setwd(strWDpath)

## process
# extract the file name from the image path
dfGC$image_path <- tools::file_path_sans_ext(basename(dfGC$image_path))
dfCO$image_path <- tools::file_path_sans_ext(basename(dfCO$image_path))
# split filename into triplet, species, managemtn, subplot and orientation
dfGC <- dfGC %>% 
  separate_wider_delim(
    image_path, delim = "_",
    names = c("triplet","species","manag","sub_plt","orient")
  )
dfCO <- dfCO %>%
  separate_wider_delim(
    image_path, delim = "_",
    names = c("triplet","species","manag","sub_plt","orient")
  )
# set datatype as factors
dfGC$triplet <- as.factor(dfGC$triplet)
dfGC$species <- as.factor(dfGC$species)
dfGC$manag <- as.factor(dfGC$manag)
dfCO$triplet <- as.factor(dfCO$triplet)
dfCO$species <- as.factor(dfCO$species)
dfCO$manag <- as.factor(dfCO$manag)

levels(dfCO$manag) <- c("cleared","dead","living")
levels(dfGC$manag) <- c("cleared","dead","living")

#### Aggregate predictions at patch level ####
dfGC_agg <- dfGC %>%
  select(!c(orient, sub_plt)) %>%
  group_by(triplet, manag, species) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  ungroup()
dfCO_agg <- dfCO %>%
  select(!c(orient, sub_plt)) %>%
  group_by(triplet, manag, species) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  ungroup()

#### Combine aggregated groundcover and complexity predictions ####
dfPrd_agg <- inner_join(
  x = dfGC_agg,
  y = dfCO_agg,
  by = join_by(triplet, species, manag)
)

#### Combine aggregated groundtruth ####
dfGroundtruth <- left_join(
  x = dfGC_og_agg,
  y = dfCO_og_agg,
  by = join_by(triplet, manag, species)
)

#### Load disturbance severity from field data ####
dfFieldSeverity <- read.csv(
  file.path(strWDpath, "R", "data_franconia", "BA_severity_estimation.csv"),
  sep = ",",
  dec = ".",
  header = TRUE
)
dfFieldSeverity$trip_n <- as.factor(dfFieldSeverity$trip_n)
dfFieldSeverity$dom_sp <- as.factor(dfFieldSeverity$dom_sp)
dfFieldSeverity[dfFieldSeverity$managed <= 0,]$managed <- 0
dfFieldSeverity[dfFieldSeverity$unmanaged <= 0,]$unmanaged <- 0

## pivot field severity longer
l_dfFieldSev <- dfFieldSeverity %>%
  select(trip_n, dom_sp, managed, unmanaged) %>%
  pivot_longer(
    cols = c(managed, unmanaged),
    names_to = "manag",
    values_to = "severity"
  )
l_dfFieldSev[l_dfFieldSev$manag == "managed",]$manag <- "cleared"
l_dfFieldSev[l_dfFieldSev$manag == "unmanaged",]$manag <- "dead"
l_dfFieldSev$manag <- factor(
  l_dfFieldSev$manag,
  levels = c("cleared", "dead", "living")
)
l_dfFieldSev$severity <- abs(l_dfFieldSev$severity)

rm(dfCO_agg, dfGC_agg, dfCO_og_agg, dfGC_og_agg)
```

| Dataframe | Description |
|------------------------------------|------------------------------------|
| dfCO | Predictions of AI/DNN about complexity of surrounding images. |
| dfCO_KH | additional complexity groundtruth |
| dfCO_og | Complexity groundtruth |
| dfFDIndis | Field data change indicators --OLD |
| dfFieldData | Contains indicators and development trajectories based on field data. |
| dfFieldSeverity | Contains estimation about disturbance severity based on field data. |
| dfGC | Predictions of AI/DNN about groundcover of plot image |
| dfGC_og | Groundcover groundtruth |
| dfGroundtruth | Contains the manually labeled complexity and groundcover values aggregated to patch level. |
| dfPrd_agg | Combination of complexity and groundcover prediction from AI, aggregated to patch level. |
| l_dfFieldSev | dfFieldSeverity in long format |

# 1. Technical validation of DNNs

Technical validation of the different Neural Network models is done in other scripts. Here the results of the 10 models for groundcover and complexity of surroundings are analysed and the best model identified.

> Code for model validation and comparison between all the models used. Currently done in Visualize_train_val.Rmd and Initial analysis.Rmd.

Training history of the models are stored in the "history.csv" file in every model folder. The results from the cross-validation runs are also of interest, they are stored in the "cross_val_res.csv" file.

```{r DNN_loadTest_result}
# Load train and validation data
dfGC_hist <- read.csv(
  file = file.path(strGcDNNPath, "history.csv"),
  header = TRUE,
  sep = ",",
  dec = "."
)
dfCO_hist <- read.csv(
  file = file.path(strCoDNNPath, "history.csv"),
  header = TRUE,
  sep = ",",
  dec = "."
)

dfGC_hist
dfCO_hist
```

The \_hist dataframes contain the 5 metrics from test and validation runs.

## 1.1 Metrics from test runs - TOCHECK

5 Metrics are evaluated: Loss, Kullback_leibler divergence, mean relative error, mean squared error and accuracy. Stored in test_res.txt.
```{r}
# Xception Model
tst_gc <- read_csv(file = file.path(strGcDNNPath, "test_res.txt"))
tst_co <- read_csv(file = file.path(strCoDNNPath, "test_res.txt"))

txt <- readLines(file.path(strGcDNNPath, "test_res.txt"))
val_score_gc <- gsub("\\[|\\]", "", txt)[3]

txt <- readLines(file.path(strCoDNNPath, "test_res.txt"))
val_score_co <- gsub("\\[|\\]", "", txt)[3]

df_tst <- bind_rows(
  tst_gc %>% mutate(src = "groundcover"),
  tst_co %>% mutate(src = "context")
)

df_tst <- df_tst %>% 
  select(
    src,
    Loss = `['loss'`,
    `KL-divergence` = `'kullback_leibler_divergence'`,
    MRE = `'mean_relative_error'`,
    MSE = `'mean_squared_error'`,
    Accuracy = `'accuracy']`
  ) %>% 
  mutate(across(c(Loss, Accuracy), ~str_remove_all(., "\\[|\\]"))) %>% 
  mutate(across(where(is.character), type.convert, as.is = TRUE)) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)))

rm(tst_co, tst_gc, txt, val_score_co, val_score_gc)

# flex table
ft <- df_tst %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(df_tst), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(df_tst), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")

ft <- fit_to_width(ft, max_width = 6.5)
ft
```

| Date | Model | Loss | Kullback leibler divergence | Mean relative error | Mean squared error | Accuracy |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| 10.05.2025 | Xception - GC | 0.02870 | 0.79323 | 0.09704 | 0.02870 | 0.50427 |
| 09.05.2025 | B0 - GC | 0.042934 | 1.21918 | 0.12814 | 0.04293 | 0.21368 |
| 09.05.2025 | B2 - GC | 0.04131 | 1.21178 | 0.12745 | 0.04131 | 0.45299 |
| 09.05.2025 | B4 - GC | 0.03479 | 1.07136 | 0.11514 | 0.03479 | 0.59829 |
| 09.05.2025 | B5 - GC | 0.03009 | 0.90542 | 0.10750 | 0.03009 | 0.51282 |
| 10.05.2025 | CNT - GC | 0.03446 | 1.19679 | 0.11466 | 0.03447 | 0.60684 |
| 10.05.2025 | CNS - GC | 0.02989 | 0.81516 | 0.10408 | 0.02989 | 0.60684 |
| 10.05.2025 | CNB - GC | 0.02883 | 0.86879 | 0.09778 | 0.02883 | 0.60684 |
| 13.05.2025 | CNL - GC | 0.03269 | 0.86643 | 0.10849 | 0.03269 | 0.60684 |
| 13.05.2025 | CNXL - GC | 0.03229 | 0.96562 | 0.10678 | 0.03229 | 0.60684 |
| 10.05.2025 | Xception - CO | 2.68000 | 0.38602 | 1.20238 | 2.68000 | 0.36842 |
| 10.05.2025 | B0 - CO | 4.82283 | 17.30694 | 1.52768 | 4.82283 | 0.22368 |
| 10.05.2025 | B2 - CO | 2.66929 | 0.17052 | 1.24081 | 2.66929 | 0.28947 |
| 10.05.2025 | B4 - CO | 4.32863 | 14.82100 | 1.42636 | 4.32863 | 0.26316 |
| 10.05.2025 | B5 - CO | 3.73054 | 8.33829 | 1.30177 | 3.73054 | 0.34211 |
| 10.05.2025 | CNT - CO | 2.56407 | -0.15456 | 1.24103 | 2.56407 | 0.32895 |
| 10.05.2025 | CNS - CO | 2.49027 | -0.17696 | 1.23825 | 2.49027 | 0.32895 |
| 10.05.2025 | CNB - CO | 2.51295 | 0.11058 | 1.23616 | 2.51295 | 0.30263 |
| 15.05.2025 | CNL - CO | 2.73880 | -0.02276 | 1.26981 | 2.73880 | 0.31579 |
| 13.05.2025 | CNXL - CO | 2.65121 | -0.10894 | 1.26630 | 2.65121 | 0.32895 |

## 1.2 Cross-validation metrics

10 fold cross validation hase been performed with the models, so the 5 before mentioned metrics are recorded for every of the 10 folds.

```{r}
# Load csv-files
dfGC_cv <- read.csv(
  file = file.path(strGcDNNPath, "cross_val_res.csv"),
  header = TRUE,
  sep = ",",
  dec = "."
)
dfCO_cv <- read.csv(
  file = file.path(strCoDNNPath, "cross_val_res.csv"),
  header = TRUE,
  sep = ",",
  dec = "."
)

### Groundcover
dfGC_cv_mean <- dfGC_cv %>%
  pivot_longer(cols = -Fold, names_to = "metrics", values_to = "val") %>% 
  group_by(metrics) %>% 
  summarize(mean_val = mean(val, na.rm = TRUE))

# Plot groundcover
ggplot(
  data = dfGC_cv %>%
    pivot_longer(cols = -Fold, names_to = "metrics", values_to = "val")
) +
  geom_line(
    aes(x = Fold, y = val, colour = metrics)
  ) +
  geom_hline(
    data = dfGC_cv_mean,
    aes(yintercept = mean_val),
    linetype = "dashed",
    colour = "black"
  ) +
  theme_bw() +
  facet_wrap(~metrics, scales = "free") +
  labs(title = "Cross-validation of groundcover DNN")


### Complexity
dfCO_cv_mean <- dfCO_cv %>% 
  pivot_longer(cols = -Fold, names_to = "metrics", values_to = "val") %>% 
  group_by(metrics) %>% 
  summarize(mean_val = mean(val, na.rm = TRUE))

# Plot complexity
ggplot(
  data = dfCO_cv %>% 
    pivot_longer(cols = -Fold, names_to = "metrics", values_to = "val")
) +
  geom_line(
    aes(x = Fold, y = val, colour = metrics)
  ) +
  geom_hline(
    data = dfCO_cv_mean,
    aes(yintercept = mean_val),
    linetype = "dashed",
    colour = "black"
  ) +
  theme_bw() +
  facet_wrap(~metrics, scales = "free") +
  labs(title = "Cross-validation of complexity DNN")


rm(dfGC_cv, dfGC_cv_mean, dfCO_cv, dfCO_cv_mean)
```

# Identifying important components

PCA to identify which variables are important, and looking into the relative change of variables from undisturbed to disturbed.

## PCA - Principal component analysis

The idea of a PCA is to reduce the number of dimensions needed to represent each data point, by searching for principal components. So for an initial dataset of N data points with P variables, K principal components are identified (1 ≤ K ≤ P). A PCA is only possible with numeric data.

```{r PCA}
### Principal component analysis -> Only possible with numeric data
## 1. normalizing data
dfPrd_agg_scl <- scale(dfPrd_agg %>% dplyr::select(-triplet, -manag, - species))

## 2. apply pca
dfPrd_agg.pca <- princomp(
  scale(
    dfPrd_agg_scl# %>% dplyr::select(-triplet, -manag, - species)
  )
)

## 3. visualization
fviz_eig(dfPrd_agg.pca, addlabels = TRUE, geom = "line")
fviz_pca_var(dfPrd_agg.pca)
fviz_pca_contrib(dfPrd_agg.pca, choice = "var", axes = 1:3)
fviz_pca_ind(
  dfPrd_agg.pca,
  repel = TRUE,
  addEllipses = TRUE,
  ellipse.level = 0.95,
  habillage = dfPrd_agg$manag
)
fviz_pca_ind(
  dfPrd_agg.pca,
  repel = TRUE,
  col.ind = dfPrd_agg$manag,
  addEllipses = TRUE,
  ellipse.type = "convex",
  mean.point = TRUE
)
```

-   3 Dimensions explain about 90% of the variation

-   Variables cluster in 4 directions

-   No clear patterns visible when plotting all entries in 2 dimensions

Useful resource: <https://f0nzie.github.io/machine_learning_compilation/detailed-study-of-principal-component-analysis.html>

```{r}
rm(dfPrd_agg.pca, dfPrd_agg_scl)
```

## Plot indicators
Plotting all indicators with facet on management and forest type.
```{r}
# ggplot(
#   data = dfPrd_agg %>% 
#     select(-c(starts_with("gd"), gc_mean, gc_mean_sd)) %>% 
#     pivot_longer(
#       cols = c(starts_with("grade_"), starts_with("gc_")), #matches("^(grade|gc)_"), #starts_with("g"),
#       names_to = c("prefix", "Indicator"),
#       names_pattern = "^(grade|gc)_(.+)$",
#       values_to = "val"
#     )
# ) +
#   theme_bw() +
#   geom_boxplot(
#     aes(x = Indicator, y = val, colour = manag)
#   ) +
#   facet_grid(species~prefix, scales = "free") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))

base_data <- dfPrd_agg %>% 
  select(-c(starts_with("gd"))) %>% 
  pivot_longer(
    cols = c(starts_with("grade_"), starts_with("gc_")), #matches("^(grade|gc)_"), #starts_with("g"),
    names_to = c("prefix", "Indicator"),
    names_pattern = "^(grade|gc)_(.+)$",
    values_to = "val"
  )

p_grade <- ggplot(
  base_data %>% filter(prefix == "grade"),
  aes(x = Indicator, y = val, colour = manag)
) +
  theme_bw() +
  geom_boxplot() +
  facet_grid(rows = vars(species), cols = vars(prefix)) +
  labs(y = "Complexity grade (0-5)", x = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

p_gc <- ggplot(
  base_data %>% filter(prefix == "gc"),
  aes(x = Indicator, y = val, colour = manag)
) +
  theme_bw() +
  geom_boxplot() +
  facet_grid(rows = vars(species), cols = vars(prefix)) +
  labs(y = "Groundcover (%)", x = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"   # will collect from p_grade
  )

p_grade + p_gc + plot_layout(widths = c(1, 1), guides = "collect") & theme(legend.position = "bottom")
ggsave(
  filename = "output/sup_indisbox.png",
  scale = 4,
  width = 600,
  height = 600,
  units = "px"
)
```


## Change of variables from undisturbed to disturbed

Relative change over all indicators/variables that were predicted by AI is calculated. For this a function is created.

$$ relative Change= \dfrac {(final Value-initial Value)} {initial Value} $$

```{r rel_change}
f_relChange <- function(p_iniVal, p_finVal) {
  return ((p_finVal - p_iniVal) / p_iniVal) # relative change
}

f_Change <- function(p_iniVal, p_finVal) {
  return (p_finVal - p_iniVal) # change
}
```

The data needs to be modified, at the moment each undisturbed plot/site has its own entry in the data-frame. So in a first step the undisturbed sites (living) are assigned to disturbed sites (dead and cleared). Two steps are taken, first the calculation of the mean and sd for groundcover and complexity grade.

```{r calc_mean&sd}
# Only run cell once!
# Calculate mean groundcover and mean complexity
dfPrd_agg <- inner_join(
  dfPrd_agg,
  dfPrd_agg %>%
    group_by(triplet, manag, species) %>%
    dplyr::summarize(
      gd_mean = mean(grade_treespecies,
                     grade_shrubs,
                     grade_herbs,
                     grade_grass,
                     grade_moss,
                     grade_deadwood,
                     grade_layers,
                     grade_mixing,
                     trim = 0
                     ),
      gd_mean_sd = sd(c(grade_treespecies,
                        grade_shrubs,
                        grade_herbs,
                        grade_grass,
                        grade_moss,
                        grade_deadwood,
                        grade_layers,
                        grade_mixing
                        )
                      ),
      gc_mean = mean(gc_Mature_Trees,
                     gc_rejuvenation,
                     gc_shrub_layer,
                     gc_mosses,
                     gc_ferns,
                     gc_herb_layer,
                     gc_grasses,
                     gc_soil.foliage,
                     gc_rock,
                     gc_deadwood.stumps,
                     trim = 0
                     ),
      gc_mean_sd = sd(c(gc_Mature_Trees,
                        gc_rejuvenation,
                        gc_shrub_layer, gc_mosses,
                        gc_ferns, gc_herb_layer,
                        gc_grasses, gc_soil.foliage,
                        gc_rock,
                        gc_deadwood.stumps)
                      )
      ),
  by = join_by(triplet, manag, species)
)
```

### calculating relative change

Second is the actual calculation of the (relative) change.

```{r calc_(rel)_change}
# Calculate all rel Changes
dfPrd_reldiff <- dfPrd_agg %>%
  group_by(triplet, manag, species) %>%
  filter(manag != "living") %>%
  left_join(
    dfPrd_agg %>% filter(manag == "living"),
    by = join_by(triplet, species),
    suffix = c("_o", "_l")) %>%
  mutate(
    gd_diff = f_relChange(gd_mean_l, gd_mean_o),
    gc_diff = f_relChange(gc_mean_l, gc_mean_o),
    
    gc_Mature_Trees_diff = f_relChange(gc_Mature_Trees_l, gc_Mature_Trees_o),
    gc_rejuvenation_diff = f_relChange(gc_rejuvenation_l, gc_rejuvenation_o),
    gc_shrub_layer_diff = f_relChange(gc_shrub_layer_l, gc_shrub_layer_o),
    gc_mosses_diff = f_relChange(gc_mosses_l, gc_mosses_o),
    gc_ferns_diff = f_relChange(gc_ferns_l, gc_ferns_o),
    gc_herb_layer_diff = f_relChange(gc_herb_layer_l, gc_herb_layer_o),
    gc_grasses_diff = f_relChange(gc_grasses_l, gc_grasses_o),
    gc_soil.foliage_diff = f_relChange(gc_soil.foliage_l, gc_soil.foliage_o),
    gc_rock_diff = f_relChange(gc_rock_l, gc_rock_o),
    gc_deadwood.stumps_diff = f_relChange(gc_deadwood.stumps_l,
                                          gc_deadwood.stumps_o),
    
    grade_treespecies_diff = f_relChange(grade_treespecies_l,
                                         grade_treespecies_o),
    grade_shrubs_diff = f_relChange(grade_shrubs_l, grade_shrubs_o),
    grade_herbs_diff = f_relChange(grade_herbs_l, grade_herbs_o),
    grade_grass_diff = f_relChange(grade_grass_l, grade_grass_o),
    grade_moss_diff = f_relChange(grade_moss_l, grade_moss_o),
    grade_deadwood_diff = f_relChange(grade_deadwood_l, grade_deadwood_o),
    grade_layers_diff = f_relChange(grade_layers_l, grade_layers_o),
    grade_mixing_diff = f_relChange(grade_mixing_l, grade_mixing_o)
  ) %>%
  select(triplet, -manag_l, manag = manag_o, species, ends_with("_diff")) %>% 
  ungroup()

dfPrd_reldiff$diff_sum <- rowSums(
  abs(
    dfPrd_reldiff %>% select(ends_with("_diff"), -gd_diff, -gc_diff)
  )
)
dfPrd_reldiff$diff_rel <- (rowSums(
  abs(
    dfPrd_reldiff %>% select(ends_with("_diff"), -gd_diff, -gc_diff)
  )
)/18)*100

l_dfPrd_reldiff <- dfPrd_reldiff %>% pivot_longer(
  cols = ends_with("_diff"),
  names_to = "param",
  values_to = "val"
)
l_dfPrd_reldiff$param <- as.factor(l_dfPrd_reldiff$param)
l_dfPrd_reldiff$param <- factor(
  l_dfPrd_reldiff$param,
  levels = sort(levels(l_dfPrd_reldiff$param))
)
```

### Plot changes

Now some plots are created. Is it possible to see a disturbance from a relative change in indicator values?

1.  Relative change across all variables
2.  Variant, difference in mean groundcover, complexity from dead/cleared to living.
3.  Variant, here the mean values of all management options are compared to each other.

```{r plot_rel_changes}
## Plot relative Changes
ggplot(
  data = l_dfPrd_reldiff %>%
    filter(param != "gd_diff" & param != "gc_diff"),
  aes(x = param, y = val)
) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(alpha = 0.15, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = 5, col = "red4") +
  scale_y_continuous(limits = c(-1,1)) +
  coord_flip() +
  labs(title = "rel. change",
       x = "Indicator",
       y = "Change") +
  theme_pubr()

p1 <- ggplot(
  data = l_dfPrd_reldiff %>% filter(param != "gd_diff" & param != "gc_diff"),
  aes(x = param, y = val, col = manag)
) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(alpha = 0.15, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = 5, col = "red4") +
  scale_y_continuous(limits = c(-1,1)) +
  coord_flip() +
  facet_wrap(~manag) +
  labs(#title = "rel. change and management",
       x = "Variable",
       y = "Change") +
  theme_pubr()

p2 <- ggplot(
  data = l_dfPrd_reldiff %>% filter(param != "gd_diff" & param != "gc_diff"),
  aes(x = param, y = val, col = species)
) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  geom_jitter(alpha = .15, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = 5, col = "red4") +
  scale_y_continuous(limits = c(-1,1)) +
  coord_flip() +
  facet_grid(manag~species) +
  theme_pubr()

# p1 / p2

ggsave(
  filename = "output/sup_relchngmngfrst.png",
  plot = p2,
  scale = 4,
  width = 600,
  height = 600,
  units = "px"
)
```

Relative change over all variables shows, that most do not show a high relative change, that there are only a few categories from the pictures showing stronger effects. Also there are small differences visible in the strength of the change when comparing managed to unmanaged.

```{r}
rm(l_dfPrd_diff, l_dfPrd_reldiff, dfPrd_diff, dfPrd_reldiff) # dataframe w. diff
```

# Function for spliting dataset into test and train sets

Later on when training machine learing models, a split into test and training set is needed to assess the models performance.

```{r}
f_SplitIntoTestTrain <- function(dataset, part_test = 0.75, All_unique = TRUE) {
  if (All_unique == TRUE) {
    # The "simple" case, no fear of duplications
    trainIDs <- sample(
      x = seq_len(nrow(dataset)),
      size = floor(part_test * nrow(dataset))
    )
    r_train <- dataset[trainIDs,] %>% na.omit()
    r_test <- dataset[-trainIDs,] %>% na.omit()
    
    return(list(r_train, r_test))
  
  } else {
    # The "dificult" case, fear of duplications in test and train set
    # Create a group_id to later split dataset according to group
    tmp <- dataset %>%
      group_by(triplet, manag, sub_plt) %>%
      mutate(group_id = cur_group_id()) %>%
      ungroup()
    
    unique_groups <- unique(tmp$group_id)

    # Split unique group_id into train and test
    set.seed(123) # reproductability
    grp_split <- sample(unique(unique_groups))
    split_point <- floor(part_test * length(unique(unique_groups)))
    
    train_groups <- grp_split[1:split_point]
    test_groups <- grp_split[(split_point + 1):length(unique_groups)]
    
    # Assign rows to training or test set, based on group_id
    train_set <- tmp %>%
      filter(group_id %in% train_groups) %>% 
      select(-group_id) %>% 
      na.omit()
    test_set <- tmp %>% 
      filter(group_id %in% test_groups) %>% 
      select(-group_id) %>% 
      na.omit()
    
    return(list(train_set, test_set))
  }
}
```

# 2. Disturbance detection

## 2.1 Disturbed / Undisturbed

7 variables are indicators for a disturbance (grade_treespecies, grade_shrubs, grade_moss, grade_mixing, grade_herbs, grade_grass, gc_grasses). A random Forest will be trained on the plot level, so with 1241 datapoints and applied to the site level, so 120 datapoints. First a random Forest ist trained on all the variables, than on the 7 identified.

```{r}
############## All variables ##############
### Classifier to check if a picture shows disturbed or undisturbed site
# Create data frame on subplot level
dfSubPlts <- inner_join(
  x = dfCO %>% group_by(triplet, species, manag, sub_plt) %>% 
    select(-orient) %>%
    summarise(across(everything(), mean)) %>% ungroup(),
  y = dfGC %>% group_by(triplet, species, manag, sub_plt) %>% 
    select(-orient) %>% 
    summarise(across(everything(), mean)) %>% ungroup(),
  by = join_by(triplet, species, manag, sub_plt)
)

# Categorize whether a subplot is disturbed or undisturbed
dfSubPlts$disturbed <- ifelse(
  test = dfSubPlts$manag != "living",
  yes = "disturbed",
  no = "undisturbed"
)
dfSubPlts$disturbed <- as.factor(dfSubPlts$disturbed)

## train model ##
set.seed(161)
# model based on all variables
tmp <- f_SplitIntoTestTrain(dataset = dfSubPlts, All_unique = FALSE)
train_set <- tmp[[1]]
test_set <- tmp[[2]]
# model based on all variables
mDist_v <- caret::train(
  disturbed ~ .,
  data = train_set %>% select(disturbed, starts_with("g")),
  method = "rf",
  trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3)
)
# Model based on all vars
test_set$prd_cls_v <- predict(mDist_v, test_set)
nrow(test_set[test_set$disturbed == test_set$prd_cls_v,]) / nrow(test_set) # 74%
#summary(test_set)
nrow(test_set) # Amount of data points ##310
nrow(test_set[test_set$disturbed == test_set$prd_cls_v,]) # Amt. corr classific.
##231


##### Plot level #####
dfPrd_agg$disturbed <- ifelse(
  test = dfPrd_agg$manag != "living",
  yes = "disturbed",
  no = "undisturbed"
)

dfPrd_agg$prd_cls_v <- predict(mDist_v, dfPrd_agg)
nrow(dfPrd_agg[dfPrd_agg$disturbed == dfPrd_agg$prd_cls_v,]) / nrow(dfPrd_agg) #86%
dfPrd_agg$disturbed <- as.factor(dfPrd_agg$disturbed)

#summary(dfPrd_agg)
cat(
  "Accuracy on test_set: ", #68%
  round(
    nrow(test_set[test_set$disturbed == test_set$prd_cls_v,])/nrow(test_set),
    digits = 2
  )*100,
  "%\n\n",
  sep = ''
)

dfPrd_agg$match_v <- ifelse(
  test = dfPrd_agg$disturbed == dfPrd_agg$prd_cls_v,
  yes = TRUE,
  no = FALSE
)
cat(
  "Accuracy on dfPrd_agg: ", #80%
  round(
    nrow(dfPrd_agg[dfPrd_agg$disturbed == dfPrd_agg$prd_cls_v,])/nrow(dfPrd_agg),
    digits = 2
    )*100,
  "%",
  sep = ''
)
mDist_v$finalModel
mDist_v

dflvlplt <- test_set
dflvlsit <- dfPrd_agg

ggarrange(
  ncol = 2,
  nrow = 1,
  labels = "AUTO",
  common.legend = TRUE,
  
  ggplot(
    data = as.data.frame(confusionMatrix(dflvlplt$prd_cls_v, dflvlplt$disturbed)$table)
  ) +
    geom_tile(
      aes(x = Reference, y = Prediction, fill = Freq),
      color = "white"
    ) +  # Fill color based on frequency
    scale_fill_gradient(low = "white", high = "#E69F00", limits = c(0,150)) +  # Gradient from white to blue
    geom_label(
      aes(
        x = Reference,
        y = Prediction,
        label = Freq
      ),
      vjust = 1
    ) +  # Add text labels for the numbers
    theme_pubr() +
    labs(title = "Test set", x = "Actual", y = "Predicted"),
  
  ggplot(
    data = as.data.frame(confusionMatrix(dflvlsit$prd_cls_v, dflvlsit$disturbed)$table)
  ) +
    geom_tile(
      aes(x = Reference, y = Prediction, fill = Freq),
      color = "white"
    ) +  # Fill color based on frequency
    scale_fill_gradient(low = "white", high = "#E69F00", limits = c(0,150)) +  # Gradient from white to blue
    geom_label(
      aes(
        x = Reference,
        y = Prediction,
        label = Freq
      ),
      vjust = 1
    ) +  # Add text labels for the numbers
    theme_pubr() +
    labs(title = "Aggregated set", x = "Actual", y = "Predicted")
)

#rm(tmp, test_set, train_set)
```

A randomForest using all available variables (18 + target variable) yielded the best results. With a accuracy of 74 % on the test dataset and 86 % on the Plot level dataset. The model trained on the 7 varibales achieved about 6% less accuracy, with 68% on test dataset and 80% on the plot lvl dataset.

```{r}
# Plot detection of disturbed/undisturbed
ggplot(data = dfPrd_agg) +
  geom_bar(
    aes(
      x = disturbed,
      fill = match_v,
    ),
    position = "fill"
  ) +
  theme_pubr()

ggplot(data = dfPrd_agg) +
  geom_bar(
    aes(
      x = manag,
      fill = match_v
    ),
    position = "fill"
  ) +
  labs(
    y = "Percent",
    fill = "Prediction: "
  ) +
  geom_label(
    aes(
      x = manag,
      fill = match_v,
      label = scales::percent(
        after_stat(count/(sum(count)/3)),
        accuracy = 0.1
      )
    ),
    stat = "count",
    position = "fill",
    show.legend = FALSE
  ) +
  scale_x_discrete(
    "Management", labels = c("Cleared", "Dead", "Living")
  ) +
  scale_fill_paletteer_d("ggthemes::wsj_red_green", direction = -1) +
  scale_y_continuous(
    labels = scales::percent,
    breaks = scales::pretty_breaks(n = 5)
  ) +
  theme_pubr()

## Confusion matrix
dfPrd_agg$disturbed <- as.factor(dfPrd_agg$disturbed)
confusionMatrix(dfPrd_agg$prd_cls_v, dfPrd_agg$disturbed)
ggplot(
  data = as.data.frame(confusionMatrix(dfPrd_agg$prd_cls_v, dfPrd_agg$disturbed)$table)
) +
  geom_tile(
    aes(x = Reference, y = Prediction, fill = Freq),
    color = "white"
  ) +  # Fill color based on frequency
  scale_fill_gradient(low = "white", high = "blue") +  # Gradient from white to blue
  geom_label(
    aes(
      x = Reference,
      y = Prediction,
      label = Freq
    ),
    vjust = 1
  ) +  # Add text labels for the numbers
  theme_minimal() +
  labs(title = "Confusion Matrix", x = "Actual", y = "Predicted")
```

## 2.2 Calculating F1-Score

```{r}
f_calcPrecRecF1 <- function(TruePositive, FalsePositive, FalseNegative, TrueNegative) {
  n <- TruePositive + FalsePositive + FalseNegative + TrueNegative
  r_precision <- TruePositive / (TruePositive + FalsePositive)
  r_recall <- TruePositive / (TruePositive + FalseNegative)
  r_accuracy <- (TruePositive + TrueNegative) / n
  r_F1 <- 2* ((r_precision * r_recall) / (r_precision + r_recall))
  
  return(
    data.frame(
      Precision = r_precision,
      Recall = r_recall,
      Accuracy = r_accuracy,
      F1 = r_F1
    )
  )
}

#test_set:
dftest <- as.data.frame(
  confusionMatrix(
    test_set$prd_cls_v, test_set$disturbed,
    mode = "everything"
  )$table
)
### Precision
TP <- as.numeric(dftest %>%
  filter(Prediction == "disturbed" & Reference == "disturbed") %>% 
  select(Freq))
FP <- as.numeric(dftest %>%
  filter(Prediction == "disturbed" & Reference == "undisturbed") %>% 
  select(Freq))
FN <- as.numeric(dftest %>%
  filter(Prediction == "undisturbed" & Reference == "disturbed") %>% 
  select(Freq))
TN <- as.numeric(dftest %>%
  filter(Prediction == "undisturbed" & Reference == "undisturbed") %>% 
  select(Freq))

dfF1 <- f_calcPrecRecF1(TP, FP, FN, TN)
dfF1$set <- "test_set"

#aggregated_set:
dfAgg <- as.data.frame(confusionMatrix(dfPrd_agg$prd_cls_v, dfPrd_agg$disturbed,
                                       mode = "everything")$table)
TP <- as.numeric(dfAgg %>% 
  filter(Prediction == "disturbed" & Reference == "disturbed") %>% 
  select(Freq))
FP <- as.numeric(dfAgg %>%
  filter(Prediction == "disturbed" & Reference == "undisturbed") %>% 
  select(Freq))
FN <- as.numeric(dfAgg %>%
  filter(Prediction == "undisturbed" & Reference == "disturbed") %>% 
  select(Freq))
TN <- as.numeric(dfAgg %>%
  filter(Prediction == "undisturbed" & Reference == "undisturbed") %>% 
  select(Freq))

dfF1 <- dfF1 %>% add_row(
  f_calcPrecRecF1(TP, FP, FN, TN), set = "aggregat"
)
dfF1$task <- "dist_detect"
dfF1
rm(TN, TP, FN, FP, dftest, dfAgg)
```

## 2.3 Calc F1 incl. Management
Calculating the F1 score while differentiating different management approaches.
```{r}
# test_set$prd_cls_v vs. test_set$disturbed
plt_lvl <- test_set %>% 
  group_by(manag) %>% 
  summarise(
    TP = sum(prd_cls_v == "disturbed" & disturbed == "disturbed"),
    TN = sum(prd_cls_v == "undisturbed" & disturbed == "undisturbed"),
    FP = sum(prd_cls_v == "disturbed"   & disturbed == "undisturbed"),
    FN = sum(prd_cls_v == "undisturbed" & disturbed == "disturbed"),
    .groups = "drop"
  ) %>% 
  mutate(
    Accuracy     = round((TP + TN) / (TP + TN + FP + FN), 3),
    Sensitivity  = round(TP / (TP + FN), 3),
    Specificity  = round(TN / (TN + FP), 3),
    Precision    = round(TP / (TP + FP), 3),
    Recall       = round(TP / (TP + FN), 3),
    F1_score     = round(2 * TP / (2 * TP + FP + FN), 3)
  )

patch_lvl <- dfPrd_agg %>% 
  group_by(manag) %>% 
  summarise(
    TP = sum(prd_cls_v == "disturbed" & disturbed == "disturbed"),
    TN = sum(prd_cls_v == "undisturbed" & disturbed == "undisturbed"),
    FP = sum(prd_cls_v == "disturbed"   & disturbed == "undisturbed"),
    FN = sum(prd_cls_v == "undisturbed" & disturbed == "disturbed"),
    .groups = "drop"
  ) %>% 
  mutate(
    Accuracy     = round((TP + TN) / (TP + TN + FP + FN), 3),
    Sensitivity  = round(TP / (TP + FN), 3),
    Specificity  = round(TN / (TN + FP), 3),
    Precision    = round(TP / (TP + FP), 3),
    Recall       = round(TP / (TP + FN), 3),
    F1_score     = round(2 * TP / (2 * TP + FP + FN), 3)
  )

tbl2_mngmnt <- bind_rows(
  plt_lvl %>% mutate(Dataset = "Plot level test set", Task = "A: Disturbance detection", Response = "Disturbed / undisturbed"),
  patch_lvl %>% mutate(Dataset = "Patch level test set", Task = "A: Disturbance detection", Response = "Disturbed / undisturbed")
) %>% 
  select(
    Task, Dataset, Response, manag,
    Precision, Recall, Accuracy, `F1-Score` = F1_score
  )

ft <- tbl2_mngmnt %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(tbl2_mngmnt), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(tbl2_mngmnt), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")

ft <- fit_to_width(ft, max_width = 6.5)
ft

# save_as_docx(ft, path = "output/tbl2_mngmnt.docx")
```

# 3. Disturbance severity

To later evaluate the model that results form this approach several functions are needed to calculate things like Root Mean Square Error (RMSE) or Mean Absolute Error (MAE).

```{r}
f_MAE <- function(p_errorSet) {
  l_sum <- 0
  l_n <- nrow(p_errorSet)
  for (i in 1:l_n) { # Calc: Mean Absolute Error (MAE)
    l_sum <- abs(p_errorSet[i,]$sev_err) + l_sum 
  }
  r_MAE <- l_sum / l_n
  return(r_MAE)
}
```

Disturbance severity was assessed with field data, as loss/difference in basal area from reference living site to disturbed site. Here we will explore whether a random forest can be trained to predict the disturbance severity of a sub-plot, so with 5 pictures and the values of the reference site.

```{r disturbance_severity}
# Join sub plot level with severity from field data
dfSev <- left_join(
  x = dfSubPlts,
  y = l_dfFieldSev,
  by = join_by(triplet == trip_n, species == dom_sp, manag)
) %>% na.omit()

# Add reference columns from forest type to dataframe
dfSev <- left_join(
  x = dfSev,
  y = left_join(
    x = dfCO %>% 
      group_by(species, manag) %>% 
      select(-triplet, -sub_plt, -orient) %>% 
      filter(manag == "living") %>% 
      summarise(across(everything(), mean)) %>% 
      ungroup(),
    y = dfGC %>% 
      group_by(species, manag) %>% 
      select(-triplet, -sub_plt, -orient) %>% 
      filter(manag == "living") %>% 
      summarise(across(everything(), mean)) %>% 
      ungroup(),
    by = join_by(species, manag)
  ) %>% select(-manag),
  by = join_by(species),
  suffix = c("", "_ref")
)

# Calculate correlation matrix
tmp_cor <- cor(
  dfSev %>%
    select(-triplet, -species,- manag, -sub_plt, -disturbed),
  use = "complete.obs"
)
ggcorrplot(
  tmp_cor,
  type = "upper",
  outline.color = "white",
  lab = TRUE
)
rm(tmp_cor) # cleanup
```

1st exploring the existing data:

```{r dist_sev_exploration}
ggplot(data = dfSev) +
  geom_boxplot(
    aes(
      x = species,
      y = severity,
      colour = species
    ),
    outlier.shape = NA
  ) +
  geom_point(
    aes(
      x = species,
      y = severity,
      colour = species
    ),
    alpha = 0.5,
    position = position_jitterdodge(jitter.width = 0.75)
  ) +
  #facet_wrap(~manag) +
  theme_pubr()

ggplot(data = dfSev) +
  geom_density(
    aes(
      x = severity,
      fill = species
    ),
    alpha = .5,
  ) +
  stat_overlay_normal_density(
    aes(x = severity),
    colour = "red3",
    linetype = "dashed"
  ) +
  facet_wrap(~species) +
  theme_pubr()

ggplot(data = dfSev) +
  geom_density(
    aes(x = severity),
    alpha = .5
  ) +
  stat_overlay_normal_density(
    aes(x = severity),
    colour = "red3",
    linetype = "dashed"
  ) +
  theme_pubr()
```

2nd create a model to predict the disturbace severity.

```{r distsev_model}
##### Plot lvl
# Split into train and test set
sev_tmp <- f_SplitIntoTestTrain(dataset = dfSev, All_unique = FALSE)
train_set <- sev_tmp[[1]]
test_set <- sev_tmp[[2]]

set.seed(161)
# Train a model for prediction
mSevFT <- caret::train(
  severity ~ .,
  data = train_set %>% select(severity, starts_with("g")),
  method = "rf",
  trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3)
)
# Report randomForest
mSevFT$finalModel
mSevFT

# Predict on test set
test_set$prd_sev <- predict(mSevFT, test_set)

# Calculate prediction error
test_set$sev_err <- test_set$prd_sev - test_set$severity

# Save to file
# write_csv(x = test_set, file = "output/data_fig3.csv")

dfsev_patch <- dfSev %>% 
  group_by(triplet, species, manag) %>% 
  summarise(across(is.numeric, mean)) %>% 
  ungroup()

# Predict on Patch level
dfsev_patch$prd_sev <- predict(mSevFT, dfsev_patch)
# Calc prediction error
dfsev_patch$sev_err <- dfsev_patch$prd_sev - dfsev_patch$severity

# Save to file
# write_csv(x = dfsev_patch, file = "output/data_fig3_patch.csv")
```

Plotting the predictions of severity.

```{r plots}
# # Plot severity to predicted severity
ggplot(
  data = test_set,
  aes(
    x = abs(severity),
    y = prd_sev
  )
) +
  geom_abline(colour = "grey") +
  geom_point(
    aes(colour = species)
  ) +
  #geom_abline(alpha = 0.5) +
  labs(
    #title = "Severity and prd. severity",
    x = "Observed severity (%)",
    y = "Predicted severity (%)",
    col = "Species"
  ) +
  ylim(c(0,100)) +
  #plot_scheme(color("bright")(4)) +
  #scale_color_paletteer_d("wesanderson::Moonrise2") +
  scale_color_paletteer_d("MetBrewer::Kandinsky") +
  theme_pubr()

summary(lm(prd_sev ~ severity, data = test_set))

cat("Summary of prediction error(sev_err):\n")
summary(test_set$sev_err)
cat("Standard deviation: ", sd(test_set$sev_err), "\n")
cat("Root Mean Squared Error(RMSE): ", f_RMSE(test_set$sev_err), "\n")
cat("Mean Absolute Error(MAE): ", f_MAE(test_set), "\n")


ggplot(
  data = dfsev_patch,# patch_tst,
  aes(
    x = abs(severity),
    y = prd_sev
  )
) +
  geom_abline(colour = "grey") +
  geom_point(
    aes(colour = species)
  ) +
  #geom_abline(alpha = 0.5) +
  labs(
    #title = "Severity and prd. severity",
    x = "Observed severity (%)",
    y = "Predicted severity (%)",
    col = "Species"
  ) +
  ylim(c(0,100)) +
  #plot_scheme(color("bright")(4)) +
  #scale_color_paletteer_d("wesanderson::Moonrise2") +
  scale_color_paletteer_d("MetBrewer::Kandinsky") +
  theme_pubr() +
  labs(title = "Patch lvl")
```

Clean the environment:

```{r}
rm(tmp, test_set, train_set, mDist_v)
```

# 4. Post-disturbance development

A disturbed forest can develop in 4 different ways, all goes back to a change in composition and/or structure. 1. Resilience (no change) 2. Reassembly (change in composition) 3. Restructuring (change in structure) 4. Replacement (change in composition and structure)

## 4.1 Dataframes

Creation of 6 dataframes for reference levels and prediction levels. Level meaning aggregated to plot, sub-plot or image level.

```{r}
# Complexity on subplot lvl
dfCO_sp <- dfCO %>% 
  group_by(triplet, species, manag, sub_plt) %>% 
  select(-orient) %>% 
  summarise(across(everything(), mean)) %>% 
  ungroup()

# Plt lvl reference
dfPltLvl_ref <- left_join(
  x = dfCO %>% filter(manag == "living"),
  y = dfGC %>% filter(manag == "living") %>% select(-orient),
  by = join_by(triplet, species, manag, sub_plt)
) %>% group_by(triplet, species, manag) %>%
  select(-sub_plt, -orient) %>% 
  summarise(across(everything(), mean)) %>% 
  ungroup()

# dataframe sub-plot level
dfSPlvl <- inner_join(
  x = dfCO_sp %>% filter(manag != "living"),
  y = dfGC %>% filter(manag != "living"),
  by = join_by(triplet, species, manag, sub_plt)
) %>% 
  left_join(
    x = .,
    y = dfFieldData %>% select(trip_n, manag, dom_sp, R_direction),
    by = join_by(triplet == trip_n, species == dom_sp, manag)
  ) %>% select(-orient)

# dataframe single-picture level
dfImglvl <- left_join(
  x = dfCO %>% filter(manag != "living"),
  y = dfGC %>% select(-orient),
  by = join_by(triplet, species, manag, sub_plt)
) %>% 
  left_join(
    x = .,
    y = dfFieldData %>% select(trip_n, manag, dom_sp, R_direction),
    by = join_by(triplet == trip_n, species == dom_sp, manag)
  )

######################
### Sub-plot level ###
dfSP_rPlt <- left_join(
  x = dfSPlvl,
  y = dfPltLvl_ref %>% select(-manag),
  by = join_by(triplet, species),
  suffix = c("","_ref")
)
dfSP_rPlt$species <- as.factor(dfSP_rPlt$species)
dfSP_rPlt$R_direction <- as.factor(dfSP_rPlt$R_direction)

### Single picture level ###
dfImg_rPlt <- left_join(
  x = dfImglvl,
  y = dfPltLvl_ref %>% select(-manag),
  by = join_by(triplet, species),
  suffix = c("","_ref")
)
dfImg_rPlt$species <- as.factor(dfImg_rPlt$species)
dfImg_rPlt$R_direction <- as.factor(dfImg_rPlt$R_direction)

#################
### Site lvl data
dfNewDat <- inner_join(
  x = dfPrd_agg[dfPrd_agg$manag != "living",] %>% select(-contains("mean")),
  y = dfPrd_agg[dfPrd_agg$manag == "living",] %>%
    select(-manag, -species, -contains("mean")),
  by = join_by(triplet),
  suffix = c("","_ref")
) %>% inner_join(
  x = .,
  y = dfFieldData %>% select(1:2, R_direction),
  by = join_by(triplet == trip_n, manag)
)
dfNewDat$R_direction <- as.factor(dfNewDat$R_direction)
```

## 4.2 Predict function

```{r}
# function to predict value and calculate error
f_predict <- function(p_model, p_data, p_actualval = NA) {
  tmp_predval <- predict(p_model, p_data)
  
  ifelse(
    test = (length(p_actualval) == 1 & is.na(p_actualval)),
    yes = return(tmp_predval),
    no = FALSE
  )
  
  tmp_prederr <- tmp_predval - p_actualval
  
  r_df <- data.frame(prediction = tmp_predval, error = tmp_prederr)
  names(r_df)[2] <- "error"
  
  return(r_df)
}
```

## 4.4 Model with all variables

Due to changes in the way the DNNs operate the 7 variables are no longer correct and all variables need to be used.

```{r}
##### Image level ##### (dfImg_rPlt)
tmp <- f_SplitIntoTestTrain(dfImg_rPlt, All_unique = FALSE)
tmp_train <- tmp[[1]] %>% na.omit()
tmp_test <- tmp[[2]] %>% na.omit()

# Model creation
set.seed(161)
mRF <- caret::train(
  R_direction ~ .,
  data = tmp_train %>% select(R_direction, starts_with("g")) %>% na.omit(),
  method = "rf",
  trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3)
)
# predict development trajectory with test data
tmp_test$PrdTrd <- f_predict(mRF, tmp_test %>% select(-R_direction))
nrow(tmp_test[tmp_test$R_direction == tmp_test$PrdTrd,]) / nrow(tmp_test) #~80%
# predict dev tradj with other data
dfNewDat$PrdTrd <- f_predict(mRF, dfNewDat %>% select(-R_direction))
nrow(dfNewDat[dfNewDat$R_direction == dfNewDat$PrdTrd,]) / nrow(dfNewDat) #~85%

##### Sub-plot level ##### (dfSP_rPlt)
tmp <- f_SplitIntoTestTrain(dfSP_rPlt, All_unique = FALSE)
tmp_train <- tmp[[1]] %>% na.omit()
tmp_test <- tmp[[2]] %>% na.omit()

#Model creation
set.seed(161)
mRF <- caret::train(
  R_direction ~ .,
  data = tmp_train %>% select(R_direction, starts_with("g")) %>% na.omit(),
  method = "rf",
  trControl = trainControl(method = "repeatedcv", number = 10, repeats = 3)
)
mRF$finalModel
mRF
# predict development trajectory with test data
tmp_test$PrdTrd <- f_predict(mRF, tmp_test %>% select(-R_direction))
nrow(tmp_test[tmp_test$R_direction == tmp_test$PrdTrd,]) / nrow(tmp_test) #~84%
# predict dev tradj with other data
dfNewDat$PrdTrd <- f_predict(mRF, dfNewDat %>% select(-R_direction, -PrdTrd))
nrow(dfNewDat[dfNewDat$R_direction == dfNewDat$PrdTrd,]) / nrow(dfNewDat)#~87.5%
```

### confusion Matrix

***Precision***: Proportion of correct positive predictions (True Positives) out of all positive prediction made by the model. It measures the *accuracy of the positive predictions*. ***Recall*** (Sensitivity or True Positive Rate): Measures the proportion of actual positive instances that were correctly identified by the model. So the *ration of True Positives to the actual positives* (True Pos. + False Negatives). ***F1-Score***: Provides a balanced measure of precision and recall, it is derived from *the harmonic mean of precision and recall*.

```{r}
# test_set
tmp <- confusionMatrix(tmp_test$PrdTrd, tmp_test$R_direction, mode = "prec_recall")
dftmp <- as.data.frame(tmp$byClass)
dftmp <- tibble::rownames_to_column(dftmp, "class")

dfF1$class <- "disturbed"
dfF1 <- dfF1 %>% add_row(
  dftmp %>% select(class, Precision, Recall, Accuracy = `Balanced Accuracy`, F1),
  set = "test_set",
  task = "development"
)
# agg_set
tmp <- confusionMatrix(dfNewDat$PrdTrd, dfNewDat$R_direction, mode = "prec_recall")
dftmp <- as.data.frame(tmp$byClass)
dftmp <- tibble::rownames_to_column(dftmp, "class")
dfF1 <- dfF1 %>% add_row(
  dftmp %>% select(class, Precision, Recall, Accuracy = `Balanced Accuracy`, F1),
  set = "aggregat",
  task = "development"
)
# write_csv(x = dfF1, file = "output/table2.csv")
dfF1
```
#### confusion matrix w. management
Expanding the confusion matrix to incoorporate management scenarios.
```{r}
# tbl2_mngmnt

plt_lvl <- tibble()
patch_lvl <- tibble()
for (mng in levels(dfNewDat$manag)) {
  # Plot level
  tmp_mng <- tmp_test %>% filter(manag == mng)
  tmp_df <- as.data.frame(
    confusionMatrix(
      data = tmp_mng$PrdTrd,
      reference = tmp_mng$R_direction,
      mode = "prec_recall"
    )$byClass
  )
  tmp_df <- rownames_to_column(tmp_df, var = "Response") %>% 
    mutate(
      manag = mng,
      Task = "B: Classifying reorganization pathway",
      Dataset = "Plot level test set"
    ) %>% 
    select(
      Task, Dataset, Response, manag, Precision, Recall,
      Accuracy = `Balanced Accuracy`,
      `F1-Score` = F1
    )
  plt_lvl <- bind_rows(plt_lvl, tmp_df)
  
  # Patch level
  tmp_mng <- dfNewDat %>% filter(manag == mng)
  
  tmp_df <- as.data.frame(
    confusionMatrix(
      data = tmp_mng$PrdTrd,
      reference = tmp_mng$R_direction,
      mode = "prec_recall"
    )$byClass
  )
  tmp_df <- rownames_to_column(tmp_df, var = "Response") %>% 
    mutate(
      manag = mng,
      Task = "B: Classifying reorganization pathway",
      Dataset = "Patch level test set"
    ) %>% 
    select(
      Task, Dataset, Response, manag, Precision, Recall,
      Accuracy = `Balanced Accuracy`,
      `F1-Score` = F1
    )
  patch_lvl <- bind_rows(patch_lvl, tmp_df)
}
tbl2_tmp <- bind_rows(tbl2_mngmnt, plt_lvl, patch_lvl) %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)))
ft <- tbl2_tmp %>% 
  flextable() %>% 
  theme_booktabs() %>%
  flextable::color(i = seq(2, nrow(df_tst), 2), color = "black", part = "body") %>%
  flextable::bg(i = seq(2, nrow(df_tst), 2), bg = "#F9F9F9", part = "body") %>%
  flextable::hline(i = c(3,6,10,14,18,22,26)) %>% 
  flextable::align(align = "center", part = "all") %>%
  flextable::autofit() %>%
  flextable::bold(part = "header")
ft
ft <- fit_to_width(ft, max_width = 6.5)
# flextable::save_as_docx(ft, path = "output/tbl2_mngmnt.docx")
```

## Plots

Visualizing development classifications

```{r}
levels(tmp_test$species) <- c("Beech", "Oak", "Pine", "Spruce")
levels(dfNewDat$species) <- c("Beech", "Oak", "Pine", "Spruce")
# write_csv(x = tmp_test, file = "output/data_fig4_plot.csv")
# write_csv(x = dfNewDat, file = "output/data_fig4_patch.csv")

ggarrange(
  nrow = 2, ncol = 1,
  common.legend = TRUE,
  labels = c("(a)", "(b)"),
  
  # Plot lvl plot
  ggplot(data = tmp_test %>%
           select(R_direction, PrdTrd, species) %>%
           pivot_longer(cols = c(R_direction, PrdTrd), names_to = "param", values_to = "val")
         ) +
    geom_bar(
      aes(
        x = interaction(param, species),
        fill = factor(val)
      ),
      position = "fill"
    ) +
    facet_grid(~species, scales = "free_x") +
    scale_x_discrete(
      "", labels = c("Predicted", "Observed"), guide = guide_axis(angle = 45)
    ) +
    labs(
      y = "Percent",
      fill = "Reorganization pathway:",
      title = "Plot level trajectory"
    ) +
    #scale_fill_paletteer_d("MetBrewer::Egypt") +
    scale_fill_paletteer_d("MetBrewer::Kandinsky") +
    #scale_fill_paletteer_d("wesanderson::Moonrise2") +
    scale_y_continuous(labels = c("0", "25%", "50%", "75%", "100%")) +
    theme_pubr(),
  
  # Patch lvl plot
  ggplot(data = dfNewDat %>%
           select(R_direction, PrdTrd, species) %>%
           pivot_longer(cols = c(R_direction, PrdTrd), names_to = "param", values_to = "val")
         ) +
    geom_bar(
      aes(
        x = interaction(param, species),
        fill = factor(val)
      ),
      position = "fill"
    ) +
    facet_grid(~species, scales = "free_x") +
    scale_x_discrete(
      "", labels = c("Predicted", "Observed"), guide = guide_axis(angle = 45)
    ) +
    labs(
      y = "Percent",
      fill = "Reorganization pathway:",
      title = "Patch level trajectory"
    ) +
    #scale_fill_paletteer_d("MetBrewer::Egypt") +
    scale_fill_paletteer_d("MetBrewer::Kandinsky") +
    #scale_fill_paletteer_d("wesanderson::Moonrise2") +
    scale_y_continuous(labels = c("0", "25%", "50%", "75%", "100%")) +
    theme_pubr()
)
```

### Summaries - Plot lvl test set

```{r}
tmp_test$TrdMatch <- tmp_test$R_direction == tmp_test$PrdTrd

summary(tmp_test[tmp_test$species == "Beech",] %>% select(2,3, TrdMatch))
summary(tmp_test[tmp_test$species == "Oak",] %>% select(2,3, TrdMatch))
summary(tmp_test[tmp_test$species == "Pine",] %>% select(2,3, TrdMatch))
summary(tmp_test[tmp_test$species == "Spruce",] %>% select(2,3, TrdMatch))

# Percentage of true development prediction
tmp_test %>%
  select(2, 3, TrdMatch) %>%
  group_by(species) %>% 
  summarise(pct_true = mean(TrdMatch) * 100)

# Overall true predictions
tmp_test %>% 
  select(2, 3, TrdMatch) %>% 
  summarise(pct_true = mean(TrdMatch) * 100)
```

### Summaries - Patch lvl

```{r}
dfNewDat$TrdMatch <- dfNewDat$R_direction == dfNewDat$PrdTrd

summary(dfNewDat[dfNewDat$species == "Beech",] %>% select(2,3, TrdMatch))
summary(dfNewDat[dfNewDat$species == "Oak",] %>% select(2,3, TrdMatch))
summary(dfNewDat[dfNewDat$species == "Pine",] %>% select(2,3, TrdMatch))
summary(dfNewDat[dfNewDat$species == "Spruce",] %>% select(2,3, TrdMatch))

# Percentage of true development prediction
dfNewDat %>%
  select(2, 3, TrdMatch) %>%
  group_by(species) %>% 
  summarise(pct_true = mean(TrdMatch) * 100)

# Overall true predictions
dfNewDat %>% 
  select(2, 3, TrdMatch) %>% 
  summarise(pct_true = mean(TrdMatch) * 100)
```
# Cleanup
Removing variables etc. from environment.
```{r cleanup}
rm(dfCO, dfCO_hist, dfCO_KH, dfCO_og, dfCO_sp, dfGC, dfGC_hist, dfGC_og)
rm(dfF1, dfFieldData, dfFieldSeverity, dfGroundtruth, dfImg_rPlt, dfImglvl)
rm(dflvlplt, dflvlsit, dfNewDat, dfPltLvl_ref, dfPrd_agg, dfSev, dfSPlvl)
rm(dfSubPlts, dftmp, l_dfFieldSev, mRF, mSevFT, sev_tmp, tmp, tmp_test)
rm(tmp_train, strCoDNNPath, strGcDNNPath, strWDpath, dfSP_rPlt)

rm(f_calcPrecRecF1, f_Change, f_MAE, f_MSE, f_predict, f_R2, f_relChange)
rm(f_RMSE, f_SplitIntoTestTrain, get_newest_directory)
```
